---
title: "p8105_hw2v2_vas2145"
output: github_document
date: "2023-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading Libraries
```{r}
library(tidyverse)
library(readxl)
```


#Problem 1
```{r}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez"))
```

```{r}
snp = 
  read_csv(
    "data/fivethirtyeight_datasets/snp.csv",
    col_types = cols(date = col_date(format = "%m/%d/%y"))) |>
  separate(date, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    year = if_else(year > 2023, year - 100, year)) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, close) 
```


```{r}
unemployment = 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

Merging datasets
```{r merge_538}
data_538 = 
  left_join(pols, snp) |>
  left_join(x = _, y = unemployment)

str(data_538)
```

#Problem 2

Loading trash wheel data, cleaning names, and calculating `homes_powered`
```{r}
trashwheel_df =
  read_excel("data/trash_data.xlsx", sheet = "Mr. Trash Wheel", skip = 1, col_names = TRUE) |> 
  janitor::clean_names() |> 
  mutate(homes_powered = (weight_tons * 500) / 30,
    trash_wheel = 1,
    year = as.numeric(year))
```

Loading Professor Trash Wheel and Gwynnda data, cleaning, and organizing.
```{r}
prof_trashwheel_df =
  read_excel("data/trash_data.xlsx", sheet = "Professor Trash Wheel", skip = 1, col_names = TRUE) |> 
  janitor::clean_names() |> 
  mutate(homes_powered = (weight_tons * 500) / 30,
         trash_wheel = 2)

gwynnda_trashwheel_df =
  read_excel("data/trash_data.xlsx", sheet = "Gwynnda Trash Wheel", skip = 1, col_names = TRUE) |> 
  janitor::clean_names() |> 
  mutate(homes_powered = (weight_tons * 500) / 30,
         trash_wheel = 3)
```


Joining the datasets:
```{r}
trashwheel_df =
  bind_rows(trashwheel_df, prof_trashwheel_df, gwynnda_trashwheel_df) |> 
  janitor::clean_names()
```

The `trashwheel_df` has `trashwheel_df observations` and `trashwheel_df col` and includes information on the trashwheel being used, the `dumpster` it places its trash in, and the amount of `homes_powered` by the trash that comes from each trashwheel. 


#Problem 3

Importing, cleaning, and tidying dataset of baseline demographics.

```{r}
baseline_df =
  read_csv("data/data_mci/MCI_baseline.csv", skip = 1) |> 
  janitor::clean_names() |> 
  mutate(
    sex = recode(sex, "1" = "male", "0" = "female"),
    apoe4 = recode(apoe4, "1" = "APOE4 carrier", "0" = "APOE4 non-carrier")) |> 
  filter(age_at_onset >= current_age | age_at_onset == ".")
```


