---
title: "p8105_hw2v2_vas2145"
output: github_document
date: "2023-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading Libraries
```{r}
library(tidyverse)
library(readxl)
```


#Problem 1
```{r}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols = 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez"))
```

```{r}
snp = 
  read_csv(
    "data/fivethirtyeight_datasets/snp.csv",
    col_types = cols(date = col_date(format = "%m/%d/%y"))) |>
  separate(date, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    year = if_else(year > 2023, year - 100, year)) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, close) 
```


```{r}
unemployment = 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

Merging datasets
```{r merge_538}
data_538 = 
  left_join(pols, snp) |>
  left_join(x = _, y = unemployment)

str(data_538)
```

#Problem 2

Loading trash wheel data, cleaning names, and calculating `homes_powered`
```{r}
trashwheel_df =
  read_excel("data/trash_data.xlsx", sheet = "Mr. Trash Wheel", skip = 1, col_names = TRUE) |> 
  janitor::clean_names() |> 
  mutate(homes_powered = (weight_tons * 500) / 30,
    trash_wheel = 1,
    year = as.numeric(year))
```

Loading Professor Trash Wheel and Gwynnda data, cleaning, and organizing.
```{r}
prof_trashwheel_df =
  read_excel("data/trash_data.xlsx", sheet = "Professor Trash Wheel", skip = 1, col_names = TRUE) |> 
  janitor::clean_names() |> 
  mutate(homes_powered = (weight_tons * 500) / 30,
         trash_wheel = 2)

gwynnda_trashwheel_df =
  read_excel("data/trash_data.xlsx", sheet = "Gwynnda Trash Wheel", skip = 1, col_names = TRUE) |> 
  janitor::clean_names() |> 
  mutate(homes_powered = (weight_tons * 500) / 30,
         trash_wheel = 3)
```


Joining the datasets:
```{r}
trashwheel_df =
  bind_rows(trashwheel_df, prof_trashwheel_df, gwynnda_trashwheel_df) |> 
  janitor::clean_names()
```

The `trashwheel_df` has `trashwheel_df observations` and `trashwheel_df col` and includes information on the trashwheel being used, the `dumpster` it places its trash in, and the amount of `homes_powered` by the trash that comes from each trashwheel. 


#Problem 3

Importing, cleaning, and tidying dataset of baseline demographics.

```{r}
baseline_df =
  read_csv("data/data_mci/MCI_baseline.csv", skip = 1) |> 
  janitor::clean_names() |> 
  mutate(
    sex = recode(sex, "1" = "male", "0" = "female"),
    apoe4 = recode(apoe4, "1" = "APOE4 carrier", "0" = "APOE4 non-carrier")) |> 
  filter(age_at_onset >= current_age | age_at_onset == ".")

developedmci_df =
  baseline_df |> 
  filter(age_at_onset != ".")
```

We first import the data required to create the the dataframe. We need to `skip = 1` since the first row does not contain data, and recode `sex` and `apoe4` so that they are appropriately encoded (not numeric). Finally, we use `filter` to remove anyone who does not have MCI at baseline. There are `baseline_df, nrow()` observations, of whom 94 develop MCI. The mean baseline age is `baseline_df, mean(pull(current_age).` The proportion of women in the study who are APOE3 carriers is 30%. 


Cleaning second MCI dataset:
```{r}
amyloid_df =
  read_csv("data/data_mci/mci_amyloid.csv", skip = 1) |> 
  janitor::clean_names() |> 
  rename(id = study_id)
```

We first import the data required to create the the amyloid dataframe. We need to `skip = 1` since the first row does not contain data, and `rename` `study_id` to `id` so that we can merge this dataset with the baseline dataset. The `amyloid_df` includes `amyloid_df, nrow()` observations.

There are 7 more participants in the amyloid dataset than in the baseline dataset. 232 of the participants are apoe4 carriers.

Combining datasets:
```{r}
combined_mci_df =
  inner_join(baseline_df, amyloid_df, by = "id")

write.csv(combined_mci_df, file = "data/data_mci/combined_mci_df.csv")

filter(combined_mci_df, sex == "female", TRUE)
filter(combined_mci_df, apoe4 == "APOE4 carrier", TRUE)

```

The combined_mci_df has 472 observations with 11 variables. 205 of participants in the combined dataset are women. 